---
title: "DATA CREW :: Project Facebook Data set Analysis"
team: "DATA CREW"
author: "JITEN / ERIK"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Libraries

```{r Imports , message=FALSE}

library(psych)
library(corrplot)
library(QuantPsyc)
library(car)
library(ggplot2)
library(ggpubr)  

```

#### Import text file

Column reference:
<ul>
<li> PTLike <- Page Total Likes
<li> Type
<li> Category (1,2,3 - product / action / inspiration)
<li> PosMon <- Post Month
<li> PosWkDay <- Post Weekday
<li> PosHr <- Post Hour
<li> Paid
<li> LPTReach <- Lifetime Post Total Reach
<li> LPTImpr <- Lifetime Post Total Impressions 
<li> LEngUser <- Lifetime Engaged Users
<li> LPConsumer <- Lifetime Post Consumers
<li> LPConsump <- Lifetime Post Consumption
<li> LPIPepLkPage <- Lifetime Post Impressions by people who have liked your Page
<li> LPRchPepLKPage <- Lifetime Post reach by people who like your Page
<li> LPepLkEngPos <- Lifetime People who have liked your Page and engaged with your post
<li> comment
<li> like
<li> share
<li> TotalInterac <- Total Interactions 
</ul>

```{r Project_read_csv }

# read the csv file
myd <- read.csv("dataset_Facebook.csv",sep = ";", header = T)

# new columns name - refer to columns reference for full description
colnames(myd) <- c('PTLike','Type','Category','PosMon','PosWkDay','PosHr','Paid',
                   'LPTReach','LPTImpr','LEngUser','LPConsumer','LPConsump',
                   'LPIPepLkPage','LPRchPepLKPage','LPepLkEngPos',
                   'comment','like','share','TotalInterac')

# display dataframe
str(myd)

```

#### Data Cleaning / Wrangling

```{r Project_data_cleaning }

# remove null/na values
myd <- na.omit(myd)

# removing features used for evaluating post impact
# and other not required variables
myd <- myd[,-c(8:10)]
myd <- myd[,-c(9:12)]
myd <- myd[,-c(9:11)]

# makes a copy of every variable in myd
attach(myd)

# create dummy variables
# Type (Photos,Status,Video,Link)
# category Factor: {action, product, inspiration }
myd$typeP=(Type=="Photo")*1
myd$typeS=(Type=="Status")*1
myd$typeV=(Type=="Video")*1
myd$category1=(Category==1)*1
myd$category2=(Category==2)*1

# remove a copy of every variable in myd
detach(myd)

# remove the column for which we created dummy variables
# also removing comment,like,share since we have total interaction
mydata <- myd[,-c(2:3)]

```

#### Explore Data

```{r Project_explore_data }

# describe the distribution of Life Time post consumers
describe(mydata$LPConsumer)

# describe the distribution of Page total likes
describe(mydata$PTLike)

# describe the distribution of totalInteraction
describe(mydata$TotalInterac)

```

```{r plot_data, message=FALSE}

# plot a histogram on the Y variable
plot_hist_lpcf <- ggplot(mydata, aes(x=LPConsumer)) + 
                  geom_histogram(bins = 30, color="black", fill="lightblue") +
                  geom_vline(aes(xintercept=mean(LPConsumer)), col="darkblue") +
                  labs(title = "Lifetime Post Consumers \n Histogram", 
                       x = "Lifetime Post Consumers", y = 'Frequency') +
                  # move the title text to the middle
                  theme(plot.title=element_text(hjust=0.5)) +
                  theme(text = element_text(size = 10))  +
                  theme(axis.title = element_text(size = 10)) +
                  theme(axis.text.x = element_text(angle = -45, hjust = .1))

# since the histogram looks exponential we try log of Y variable
plot_hist_lpcl <- ggplot(mydata, aes(x=log(LPConsumer))) + 
                  geom_histogram(bins = 30, color="black", fill="lightblue") +
                  geom_vline(aes(xintercept=mean(log(LPConsumer))), col="darkblue") +
                  labs(title = "Lifetime \n Post Consumers \n Log Histogram", 
                       x = "Log Lifetime Post Consumers", y = 'Frequency') +
                  # move the title text to the middle
                  theme(plot.title=element_text(hjust=0.5)) +
                  theme(text = element_text(size = 10))  +
                  theme(axis.title = element_text(size = 10)) +
                  theme(axis.text.x = element_text(angle = -45, hjust = .1))

# combine histogram plots
hist_com_plot <- ggarrange(plot_hist_lpcf, plot_hist_lpcl,  
                 labels = c("Fig A", "Fig B"),
                 font.label = list(size = 9, color = "blue"))
# plot all
hist_com_plot

# scatter plots for life time post consumers vs independent variables
plot_scatter <- ggplot(mydata, aes(x = PTLike, y = log(LPConsumer))) +
              geom_point() + 
              labs(title="Life Time Post Consumer \n vs \n Page Total Likes",
              x="Page Total Likes", y = "Lifetime Post Consumers") +
              # move the title text to the middle
              theme(plot.title=element_text(hjust=0.5)) +
              theme(text = element_text(size = 10))  +
              theme(axis.title = element_text(size = 10)) 

# scatter plots for life time post consumers vs total interaction no log 
plot_scatter_1 <- ggplot(mydata, aes(x = TotalInterac, y = log(LPConsumer))) +
              geom_point() + 
              labs(title="Life Time Post Consumer \n vs \n Total Interaction",
              x="Total Interaction", y = "Lifetime Post Consumers") +
              # move the title text to the middle
              theme(plot.title=element_text(hjust=0.5)) +
              theme(text = element_text(size = 10))  +
              theme(axis.title = element_text(size = 10))


# scatter plots for life time post consumers vs total interaction: log
plot_scatter_2 <- ggplot(mydata, aes(x = log(TotalInterac), y = log(LPConsumer))) +
              geom_point() + 
              labs(title="Life Time Post Consumer \n vs \n Total Interaction",
              x="Total Interaction", y = "Lifetime Post Consumers") +
              # move the title text to the middle
              theme(plot.title=element_text(hjust=0.5)) +
              theme(text = element_text(size = 10))  +
              theme(axis.title = element_text(size = 10)) 


# Box plot Paid
p_box_paid <- ggplot(mydata, aes(x=as.factor(Paid), y=log(LPConsumer), fill=Paid)) + 
              geom_boxplot(alpha=0.8) + labs(x="Paid", y = "Life Time Post Comsumer") +
              theme(legend.position="none") + theme(axis.title = element_text(size = 10))

# we need to do something about this box plot - too busy
p_box_hour <- ggplot(mydata, aes(x=as.factor(PosHr), y=log(LPConsumer), fill=PosHr)) + 
              geom_boxplot(alpha=0.8) + labs(x="Post Hour", y = "Life Time Post Comsumer") +
              theme(legend.position="none") + theme(axis.title = element_text(size = 10))

# maybe we need to do something about this box plot - bit busy
p_box_month <- ggplot(mydata, aes(x=as.factor(PosMon), y=log(LPConsumer), fill=PosMon)) + 
          geom_boxplot(alpha=0.8) + labs(x="Post Month", y = "Life Time Post Comsumer") +
          theme(legend.position="none") + theme(axis.title = element_text(size = 10))

# week day post
p_box_weekd <- ggplot(mydata, aes(x=as.factor(PosWkDay), y=log(LPConsumer), 
          fill=PosWkDay)) + geom_boxplot(alpha=0.8) + labs(x="Post Weekday", 
          y = "Life Time Post Comsumer") + theme(legend.position="none") +
          theme(axis.title = element_text(size = 10))

# combine histogram plots
box_com_plot <- ggarrange(p_box_paid, p_box_hour, p_box_month, p_box_weekd, 
                labels = c("Fig A", "Fig B", "Fig C", "Fig D"),
                font.label = list(size = 9, color = "blue"))

# scatter plot
plot_scatter
plot_scatter_1
plot_scatter_2

# plot all
box_com_plot

# correlation of the dataset  
mydata_cor <- cor(mydata[, names(mydata) %in% c("PTLike", "LPConsumer", "TotalInterac")], 
                  method = "pearson")

corrplot(mydata_cor,type="lower", method = 'number', addCoef.col = 'brown',
         number.cex = 0.8, tl.cex = 0.8)

# display correlation values
mydata_cor

```

### Model Building

```{r model_building }

# model including all relevant variables 
fit_full_1 <- lm(log(LPConsumer) ~ PTLike + PosHr + Paid + log(TotalInterac) + typeP + typeS +
                   typeV + category1 + category2, data=mydata)

# summary of full model
summary(fit_full_1) 

```
#sinks the data into connection as text file

```{r variable_selection }

# stepwise variable selection on the full result also gave same result 
step(fit_full_1, direction = "backward", trace = FALSE)

```

```{r finalize_fit_full_model }

# using stepwise variable selction
fit_full_2 <- lm(formula = log(LPConsumer) ~ PTLike + Paid + log(TotalInterac) + 
                   typeP + typeS + typeV + category1, data = mydata)


# summary of full model
summary(fit_full_2) 

# analysis of variance
anova(fit_full_2)


# removing variable due to t-value & F-value: paid 
fit_full_4 <- lm(formula = log(LPConsumer) ~ PTLike + log(TotalInterac) + 
                   typeP + typeS + typeV + category1, data = mydata)

# summary of full model
summary(fit_full_4) 

# analysis of variance
anova(fit_full_4)

# check multicollinearity
vif(fit_full_4)

# 95% confidence interval of the fitted model
confint(fit_full_4, level=0.95)

```

#### Influential Points and Outliers

```{r outliers_influential }

# plot of deleted studentized residuals vs hat values
student_hat <- data.frame(rstudent = rstudent(fit_full_4), hatvalues =  hatvalues(fit_full_4)) 

# plot rstudent vs hatvalues
student_hat_plot <- ggplot(student_hat, aes(x = hatvalues, y = rstudent)) + geom_point() +
                    # Change line type and color
                    geom_hline(yintercept=3, linetype="dashed", color = "red") +
                    geom_hline(yintercept=-3, linetype="dashed", color = "red")

# plot
student_hat_plot

# outliers |standardized residuals| > 3
std_residual = data.frame(residual = rstandard(fit_full_4)) 

# display |standardized residuals| > 3
filter(std_residual, abs(residual) > 3)

# historgram for outliers
ggplot(std_residual, aes(x = residual)) + 
  geom_histogram(bins=30, color="black", fill="lightblue") +
  labs(title = "Lifetime Post Consumers \n Standarized Residual", x = "Standarized Residual", 
       y = 'Frequency') +
      geom_vline(xintercept = 3, linetype="dotted", 
                color = "red") +
      geom_vline(xintercept = -3, linetype="dotted", 
                color = "red") +
      # move the title text to the middle
      theme(plot.title=element_text(hjust=0.5)) +
      theme(text = element_text(size = 10))  +
      theme(axis.title = element_text(size = 10)) 

# print out only observations that may be influential 
summary(influence.measures(fit_full_4))

# influential Plot
influencePlot(fit_full_4, scale=5, xlab="Hat-Values", ylab="Studentized Residuals", 
              fill.col=carPalette()[2], fill.alpha=0.5, id=TRUE)

```

#### First Final Model

```{r final_model_validation}

# standardized beta coefficient 
lm.beta(fit_full_4)

# residual vs fitted Model
residual_plot <-  ggplot(fit_full_4, aes(x = .fitted, y = .resid)) +
                  geom_point() +
                  geom_hline(yintercept = 0, col = "red") +
                  labs(title="Fitted",
                    x = "Fitted", y = "Residual") +
                  # move the title text to the middle
                  theme(plot.title=element_text(hjust=0.5))  +
                  theme(text = element_text(size = 10))  +
                  theme(axis.title = element_text(size = 10))

# Page Total Likes vs residuals
PTLike_plot <-  ggplot(mydata, aes(x = PTLike, 
                y = rstandard(fit_full_4))) + geom_point() +
                geom_hline(yintercept = 0, col = "red") +
                labs(title="Page Total Likes",
                x = "Page Total Likes", y = "Residual") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

#create Q-Q plot
qq_plot <-  ggplot(fit_full_4, aes(sample=rstandard(fit_full_4))) +
                stat_qq(size=1.5, color='blue') + 
                stat_qq_line(col = "red") +
                labs(title="Q-Q Plot",
                  x = "Theoretical Quantiles", y = "Sample Quantiles") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

# combine all plots
fit_final_plot <- ggarrange(residual_plot, PTLike_plot, qq_plot, 
                 labels = c("Fig A", "Fig B", "Fig C"),
                 font.label = list(size = 9, color = "blue"))
# plot all
fit_final_plot

```

#### Interaction Model Building

```{r interaction_model_building}

# model interaction of Paid with other independent variables
fit_inter_Paid_full_1 <- lm(log(LPConsumer) ~ Paid*(PTLike  + typeP + typeS + log(TotalInterac) +
                   typeV + category1) , data=mydata)

# summary Report
summary(fit_inter_Paid_full_1)

```

#### Interaction variable selection

```{r interaction_variable_selection}

# stepwise variable selection on the full interaction model
step(fit_inter_Paid_full_1, direction = "backward", trace = FALSE)

```

```{r interaction_Paid_model}

# selecting the model from stepwise backward variable selection process
fit_inter_Paid_2 <- lm(formula = log(LPConsumer) ~ Paid + PTLike + typeP + typeS + 
                log(TotalInterac) + typeV + category1, data = mydata)

# after stepwise fit model 
summary(fit_inter_Paid_2)

# removing non significant variable paid
fit_inter_Paid_3 <- lm(formula = log(LPConsumer) ~ PTLike + typeP + typeS + 
                log(TotalInterac) + typeV + category1, data = mydata)

# after stepwise fit model 
summary(fit_inter_Paid_3)

# analysis of variance
anova(fit_inter_Paid_3)


# 95% confidence interval of the fitted model
confint(fit_inter_Paid_3, level=0.95)

# influential Plot
influencePlot(fit_inter_Paid_3)

```

#### Influential Points and Outliers

```{r outliers_influential_Paid_inter }

# plot of deleted studentized residuals vs hat values
student_hat <- data.frame(rstudent = rstudent(fit_inter_Paid_3), hatvalues = 
                            hatvalues(fit_inter_Paid_3)) 

# plot rstudent vs hatvalues
student_hat_plot <- ggplot(student_hat, aes(x = hatvalues, y = rstudent)) + geom_point() +
                    # Change line type and color
                    geom_hline(yintercept=3, linetype="dashed", color = "red") +
                    geom_hline(yintercept=-3, linetype="dashed", color = "red")

# plot
student_hat_plot

# outliers |standardized residuals| > 3
std_residual = data.frame(residual = rstandard(fit_inter_Paid_3)) 

# display |standardized residuals| > 3
filter(std_residual, abs(residual) > 3)

# historgram for outliers
ggplot(std_residual, aes(x = residual)) + 
  geom_histogram(bins=30, color="black", fill="lightblue") +
  labs(title = "Lifetime Post Consumers \n Standarized Residual", x = "Standarized Residual", 
       y = 'Frequency') +
      geom_vline(xintercept = 3, linetype="dotted", 
                color = "red") +
      geom_vline(xintercept = -3, linetype="dotted", 
                color = "red") +
      # move the title text to the middle
      theme(plot.title=element_text(hjust=0.5)) +
      theme(text = element_text(size = 10))  +
      theme(axis.title = element_text(size = 10)) 

# print out only observations that may be influential 
summary(influence.measures(fit_inter_Paid_3))

# influential Plot
influencePlot(fit_inter_Paid_3, scale=5, xlab="Hat-Values", ylab="Studentized Residuals",
            fill.col=carPalette()[2], fill.alpha=0.5, id=TRUE)

```

### Model Interation Validation

```{r fit_inter_final_2_validation}

# residual vs fitted Model
residual_plot <-  ggplot(fit_inter_Paid_3, aes(x = .fitted, y = .resid)) +
                  geom_point() +
                  geom_hline(yintercept = 0, col = "red") +
                  labs(title="Fitted",
                    x = "Fitted", y = "Residual") +
                  # move the title text to the middle
                  theme(plot.title=element_text(hjust=0.5))  +
                  theme(text = element_text(size = 10))  +
                  theme(axis.title = element_text(size = 10))

# Page Total Likes vs residuals
PTLike_Paid_plot <-  ggplot(mydata, aes(x = PTLike, y =rstandard(fit_inter_Paid_3))) +
                geom_point() +
                geom_hline(yintercept = 0, col = "red") +
                labs(title="Page Total Likes",
                x = "Page Total Likes", y = "Residual") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

#create Q-Q plot
qq_plot <-  ggplot(fit_inter_Paid_3, aes(sample=rstandard(fit_inter_Paid_3))) +
                stat_qq(size=1.5, color='blue') + 
                stat_qq_line(col = "red") +
                labs(title="Q-Q Plot",
                  x = "Theoretical Quantiles", y = "Sample Quantiles") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

# combine all plots
fit_inter_Paid_plot <- ggarrange(residual_plot, PTLike_Paid_plot, qq_plot, 
                 labels = c("Fig A", "Fig B", "Fig C"),
                 font.label = list(size = 9, color = "blue"))
# plot all
fit_inter_Paid_plot

```

#### Full Interaction Model

```{r full_interaction_model}

# full interaction model
fit_full_interaction_1 <- lm(log(LPConsumer) ~ (PTLike + Paid + typeP + typeS + log(TotalInterac) +
                          typeV + category1)^2 , data=mydata)

# summary results
summary(fit_full_interaction_1)

```

#### Full Interaction Variable Selection

```{r full_interaction_variable_selection}

# stepwise variable selection on the full interaction model
step(fit_full_interaction_1, direction = "backward", trace = FALSE)

```

```{r full_interaction_varable_model}

# selecting the model from stepwise backward variable selection process
fit_full_interaction_2 <- lm(formula = log(LPConsumer) ~ PTLike + Paid + typeP + typeS + 
                          log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) + Paid:typeP + Paid:log(TotalInterac) +
                          typeP:log(TotalInterac) +
                          log(TotalInterac):category1, data = mydata)

# after stepwise fit model 
summary(fit_full_interaction_2)

# analysis of variance
anova(fit_full_interaction_2)


# remove interaction variable due to F-value: Paid:typeP 
fit_full_interaction_3 <- lm(formula = log(LPConsumer) ~ PTLike + Paid + typeP + typeS + 
                          log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) + Paid:log(TotalInterac) +
                          typeP:log(TotalInterac) +
                          log(TotalInterac):category1, data = mydata)

# summary model display 
summary(fit_full_interaction_3)

# analysis of variance
anova(fit_full_interaction_3)


# remove interaction variable due to t-value: log(TotalInterac):category1
fit_full_interaction_3_1 <- lm(formula = log(LPConsumer) ~ PTLike + Paid + typeP + typeS + 
                        log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) + Paid:log(TotalInterac) +
                          typeP:log(TotalInterac) , data = mydata)

# summary model display 
summary(fit_full_interaction_3_1)

# remove interaction variable due to t-value: Paid:log(TotalInterac)
fit_full_interaction_3_2 <- lm(formula = log(LPConsumer) ~ PTLike + Paid + typeP + typeS + 
                          log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) +
                          typeP:log(TotalInterac) , data = mydata)

# summary model display 
summary(fit_full_interaction_3_2)

# remove interaction variable due to t-value: Paid since interaction is gone
fit_full_interaction_4 <- lm(formula = log(LPConsumer) ~ PTLike + typeP + typeS + 
                          log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) +
                          typeP:log(TotalInterac) , data = mydata)

# summary model display 
summary(fit_full_interaction_4)

# equation
equatiomatic::extract_eq(fit_full_interaction_4, use_coefs = TRUE)

# analysis of variance
anova(fit_full_interaction_4)

# 95% confidence interval of the fitted model
confint(fit_full_interaction_4, level=0.95)

# influential Plot
influencePlot(fit_full_interaction_4)

lm.beta(fit_full_interaction_4)

```

#### Full Interaction Influential Points and Outliers

```{r outliers_influential_Full_inter }

# plot of deleted studentized residuals vs hat values
student_hat <- data.frame(rstudent = rstudent(fit_full_interaction_4), hatvalues = 
                            hatvalues(fit_full_interaction_4)) 

# plot rstudent vs hatvalues
student_hat_plot <- ggplot(student_hat, aes(x = hatvalues, y = rstudent)) + geom_point() +
                    # Change line type and color
                    geom_hline(yintercept=3, linetype="dashed", color = "red") +
                    geom_hline(yintercept=-3, linetype="dashed", color = "red")

# plot
student_hat_plot

# outliers |standardized residuals| > 3
std_residual = data.frame(residual = rstandard(fit_full_interaction_4)) 

# display |standardized residuals| > 3
filter(std_residual, abs(residual) > 3)

# historgram for outliers
ggplot(std_residual, aes(x = residual)) + 
  geom_histogram(bins=30, color="black", fill="lightblue") +
  labs(title = "Lifetime Post Consumers \n Standarized Residual", x = "Standarized Residual", 
       y = 'Frequency') +
      geom_vline(xintercept = 3, linetype="dotted", 
                color = "red") +
      geom_vline(xintercept = -3, linetype="dotted", 
                color = "red") +
      # move the title text to the middle
      theme(plot.title=element_text(hjust=0.5)) +
      theme(text = element_text(size = 10))  +
      theme(axis.title = element_text(size = 10)) 

# print out only observations that may be influential 
summary(influence.measures(fit_full_interaction_4))

# influential Plot
influencePlot(fit_full_interaction_4, scale=5, xlab="Hat-Values", 
              ylab="Studentized Residuals",
              fill.col=carPalette()[2], fill.alpha=0.5, id=TRUE)

```

### Full Interaction Model Validation

```{r fit_full_inter_final_4_validation}

# residual vs fitted Model
residual_plot <-  ggplot(fit_full_interaction_4, aes(x = .fitted, y = .resid)) +
                  geom_point() +
                  geom_hline(yintercept = 0, col = "red") +
                  labs(title="Fitted",
                    x = "Fitted", y = "Residual") +
                  # move the title text to the middle
                  theme(plot.title=element_text(hjust=0.5))  +
                  theme(text = element_text(size = 10))  +
                  theme(axis.title = element_text(size = 10))

# Page Total Likes vs residuals
PTLike_Full_plot <- ggplot(mydata, aes(x = PTLike, y =rstandard(fit_full_interaction_4))) +
                geom_point() +
                geom_hline(yintercept = 0, col = "red") +
                labs(title="Page Total Likes",
                x = "Page Total Likes", y = "Residual") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

#create Q-Q plot
qq_plot <- ggplot(fit_full_interaction_4, aes(sample=rstandard(fit_full_interaction_4))) +
                stat_qq(size=1.5, color='blue') + 
                stat_qq_line(col = "red") +
                labs(title="Q-Q Plot",
                  x = "Theoretical Quantiles", y = "Sample Quantiles") +
                # move the title text to the middle
                theme(plot.title=element_text(hjust=0.5))  +
                theme(text = element_text(size = 10))  +
                theme(axis.title = element_text(size = 10))

# combine all plots
fit_inter_Paid_plot <- ggarrange(residual_plot, PTLike_Paid_plot, qq_plot, 
                 labels = c("Fig A", "Fig B", "Fig C"),
                 font.label = list(size = 9, color = "blue"))
# plot all
fit_inter_Paid_plot

```

#### Model Validation

```{r model_validation_train_test}

# setting the seed to get the same result when knit
set.seed(2500)

# split samples (75% for training and 25% for testing)
select.mydata <- sample(1:nrow(mydata), 0.75*nrow(mydata))

# selecting 75% of the data for training purpose
train.mydata <- mydata[select.mydata,]

# selecting 25% (remaining) of the data for testing
test.mydata <- mydata[-select.mydata,]


# Model: 1 : fit_full_4
fit_m1_trn <- lm(formula = log(LPConsumer) ~ PTLike + log(TotalInterac) + 
                   typeP + typeS + typeV + category1, data = train.mydata)

# summary of fit_m1_trn
summary(fit_m1_trn)

# create fitted values using test.mydata
y_pred <- predict.glm(fit_m1_trn,test.mydata)
y_obs  <- log(test.mydata[,"LPConsumer"])

# validation statistics
# RMSE of prediction error
rmse_m1 <-sqrt((y_obs-y_pred)%*%(y_obs-y_pred)/nrow(test.mydata))

# compute MAE
mae_m1 <- mean(abs(y_obs-y_pred))

# compute MAPE
mape_m1 <- mean(abs((y_obs-y_pred)/y_obs))*100

# compute cross-validated R^2_pred
r2_pred  <-  cor(cbind(y_obs,y_pred))**2
r2_train <-  summary(fit_m1_trn)$r.squared
diffr2_m1 <- abs(r2_train-r2_pred)

# print difference of cross-validate R2 and R2
# diffr2_m1[1,2]


 # Model: 2 : fit_inter_Paid_3
 fit_int_m1_trn <- lm(formula = log(LPConsumer) ~ PTLike + typeP + typeS + 
                log(TotalInterac) + typeV + category1, data = train.mydata)

 # summary of fit_int_m1_trn
 summary(fit_int_m1_trn)
#
# create fitted values using test.mydata
y_pred2 <- predict.glm(fit_int_m1_trn,test.mydata)
y_obs2  <- log(test.mydata[,"LPConsumer"])

# # validation statistics
# # RMSE of prediction error
rmse_m1_2 <-sqrt((y_obs2-y_pred2)%*%(y_obs2-y_pred2)/nrow(test.mydata))

# compute MAE
mae_m1_2 <- mean(abs(y_obs2-y_pred2))

# compute MAPE
mape_m1_2 <- mean(abs((y_obs2-y_pred2)/y_obs2))*100
#
# compute cross-validated R^2_pred
r2_pred2    <-  cor(cbind(y_obs2,y_pred2))**2
r2_train2   <-  summary(fit_int_m1_trn)$r.squared
diffr2_m1_2 <- abs(r2_train2-r2_pred2)


# print difference of cross-validate R2 and R2
# diffr2_m1_2[1,2]


# Model 3 : fit_full_interaction_4
fit_int_m2_trn <- lm(formula = log(LPConsumer) ~ PTLike + typeP + typeS + 
                          log(TotalInterac) + typeV + category1 + PTLike:typeP + 
                          PTLike:log(TotalInterac) +
                          typeP:log(TotalInterac), data = train.mydata)

# summary of fit_full_interaction_4
summary(fit_int_m2_trn)

# create fitted values using test.mydata
y_pred3 <- predict.glm(fit_int_m2_trn,test.mydata)
y_obs3  <- log(test.mydata[,"LPConsumer"])

# validation statistics
# RMSE of prediction error
rmse_m3 <- sqrt((y_obs3-y_pred3)%*%(y_obs3-y_pred3)/nrow(test.mydata))

# compute MAE
mae_m3 <- mean(abs(y_obs3-y_pred3))

# compute MAPE
mape_m3 <- mean(abs((y_obs3-y_pred3)/y_obs3))*100

# compute cross-validated R^2_pred
r2_pred3  <- cor(cbind(y_obs3,y_pred3))**2
r2_train3 <- summary(fit_int_m2_trn)$r.squared
#diffr2_m3 <- abs(r2_train3-r2_pred3)

# print difference of cross-validate R2 and R2
# diffr2_m3[1,2]

# create dataframe
Model   <-  c("fit_m1_trn", "fit_inter_Paid_3","fit_full_interaction_4")
RMSE    <-  c(rmse_m1, rmse_m1_2, rmse_m3)
MAE     <-  c(mae_m1, mae_m1_2, mae_m3)
MAPE    <-  c(mape_m1, mape_m1_2, mape_m3)
#Diff_R2 <-  c(diffr2_m1[1,2], diffr2_m1_2[1,2], diffr2_m3[1,2])

df <- data.frame(Model, RMSE, MAE, MAPE)

# print Model Info
df


#                   Model      RMSE       MAE     MAPE
#1             fit_m1_trn 0.5048645 0.3842943 6.209016
#2       fit_inter_Paid_3 0.5048645 0.3842943 6.209016
#3 fit_full_interaction_4 0.4768748 0.3701487 6.000214

# Model 3 (fit_full_interaction_4) minimizes all three validation matrics and we can conclude that it provides more accurate prediction which is closer to actual values
```